---
title: "USDA"
author: "Cynthia O'Donnell"
date: "06/29/2015"
output: html_document
---

# USDA Food Analysis

### Load and View Data

```{r}
USDA = read.csv('USDA.csv')

str(USDA)

summary(USDA)
```

### Explore Variables

What is the food with the maximum sodium count?

```{r, echo=FALSE}
which.max(USDA$Sodium)
```

How do we find out what food this is?

```{r}
names(USDA)

USDA$Description[265]
```

Oh, it's Salt.  Well, that makes sense.

```{r}
HighSodium = subset(USDA, Sodium>10000)

nrow(HighSodium)
```

There are 10 foods that are high sodium.  WHich ones are they?

```{r}
HighSodium$Description
```

I thought caviar would be high sodium.  Let's figure out where that is.

```{r}
match("CAVIAR", USDA$Description)
USDA$Sodium[4154]
```

Actually, you can do this in one line.

```{r}
USDA$Sodium[match("CAVIAR", USDA$Description)]
```

### Explore Summary Statistics

```{r}
summary(USDA$Sodium)

sd(USDA$Sodium)
```

NA?  Let's fix that!

```{r}
sd(USDA$Sodium, na.rm=TRUE)
```

Now we know that the mean sodium level of all foods in this dataset is 322.1 and the standard deviation is 1045.417.  Caviar, with a sodium count of 1500 is above one standard deviation from the mean sodium count.

### Creating Basic Plots

```{r}
plot(USDA$Protein, USDA$TotalFat)
```

Let's pretty this up a little bit.

```{r}
plot(USDA$Protein, USDA$TotalFat, xlab="Protein", ylab="Fat", main="Protein vs Fat", col="red")
```

```{r}
hist(USDA$VitaminC, xlab="Vitamin C (mg)", main="Histogram of Vitamin C Levels")
```

Hm...Let's change the x-axis, add more breaks and adjust the bin size.

```{r}
hist(USDA$VitaminC, xlab="Vitamin C (mg)", 
     main="Histogram of Vitamin C Levels", 
     xlim=(c(0, 100)),
     breaks=2000)
```

```{r}
boxplot(USDA$Sugar, main="Boxplot of Sugar Levels", ylab="Sugar (g)")
```

### Adding Variables to a Data Frame

```{r}
USDA$Sodium[1] > mean(USDA$Sodium, na.rm=TRUE)
USDA$Sodium[50] > mean(USDA$Sodium, na.rm=TRUE)

HighSodium = USDA$Sodium > mean(USDA$Sodium, na.rm=TRUE)
str(HighSodium)
```

We can change the data type to numeric instead of logical.

```{r}
HighSodium = as.numeric(HighSodium)
str(HighSodium)
```

Now we can add HighSodium to the dataframe.
```{r}
USDA$HighSodium = HighSodium
```

Let's add HighFat, High Protein and HighSugar to the dataframe.

```{r}
USDA$HighProtein = as.numeric(USDA$Protein > mean(USDA$Protein, 
                                                  na.rm=TRUE))
USDA$HighFat = as.numeric(USDA$TotalFat > mean(USDA$TotalFat, 
                                               na.rm=TRUE))
USDA$HighSugar = as.numeric(USDA$Sugar > mean(USDA$Sugar, 
                                              na.rm=TRUE))
```

### Summary Tables

```{r}
table(USDA$HighSodium)
```

This explains that 4884 foods have less sodium that the mean sodium level and 2090 foods have above average sodium levels.

```{r}
table(USDA$HighSodium, USDA$HighFat)
```

The rows correspond to the first input High Sodium and the columns correspond to the second input, High Fat.

So we have 3529 low sodium, low fat foods; 1355 low sodium, high fat foods; 1378 low fat, high sodium foods and 712 high sodium, high fat foods.

***

### Using tapply

```{r}
tapply(USDA$Iron, USDA$HighProtein, mean, na.rm=TRUE)
```

Here we group the first parameter by the second parameter and then apply the third parameter to the new grouping.

So we have mean iron counts of 2.55 in low protein foods and 3.197 in high protein foods.

```{r}
tapply(USDA$VitaminC, USDA$HighSugar, max, na.rm=TRUE)
```

Here we group the Vitamin C variable by whether or not it is high sugar, then we find the max of each group.

```{r}
tapply(USDA$VitaminC, USDA$HighSugar, summary, na.rm=TRUE)
```

This shows us the summary of the Vitamin C level of the two groups.









