---
title: "MoneyBall"
author: "Cynthia O'Donnell"
date: "06/29/2015"
output: html_document
---

#Apply MoneyBall Techniques to Predict Making NBA Playoffs

```{r, echo=FALSE}
NBA = read.csv('NBA_train.csv')
str(NBA)
```

Let's see how number of wins relates to making the playoffs.

```{r, echo=FALSE}
# How many wins to make the playoffs?
table(NBA$W, NBA$Playoffs)
```

Looking at this output we can see that teams that win fewer than 35 games almost never make it to the playoffs, while teams that wine more than 45 games almost always make it to the playoffs.

We will build a model using the difference between points scored and points allowed to predict the number of games a team will win.

We start with adding the Points Diff variable to the dataframe.

```{r}
# Compute Points Difference
NBA$PTSdiff = NBA$PTS - NBA$oppPTS

# Check for linear relationship
plot(NBA$PTSdiff, NBA$W)

# Linear regression model for wins
WinsReg = lm(W ~ PTSdiff, data=NBA)
summary(WinsReg)
```

If a team needs to win 42 games to make it to the playoffs, we would solve the equation given to us by this model.

The equation is:

Wins = 41 + 0.0326 * PTSdiff

So:

PTSdiff >= (42 - 41) / 0.0326

PTSdiff >= 30.67

The team needs to score 31 more points than it gives up to make it to the playoffs.

Now, we will use other statistics to predict PTSdiff.

```{r}
# Linear regression model for points scored
PointsReg = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + DRB + TOV + STL + BLK, data=NBA)
summary(PointsReg)

# Sum of Squared Errors

plot(PointsReg$residuals)

SSE = sum(PointsReg$residuals^2)
SSE

# Root mean squared error
RMSE = sqrt(SSE/nrow(NBA))
RMSE

# Average number of points in a season
mean(NBA$PTS)

# Remove insignifcant variables
summary(PointsReg)

PointsReg2 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + DRB + STL + BLK, data=NBA)
summary(PointsReg2)

plot(PointsReg2$residuals)

PointsReg3 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + STL + BLK, data=NBA)
summary(PointsReg3)

plot(PointsReg3$residuals)

PointsReg4 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + STL, data=NBA)
summary(PointsReg4)

plot(PointsReg4$residuals)

# Compute SSE and RMSE for new model
SSE_4 = sum(PointsReg4$residuals^2)
RMSE_4 = sqrt(SSE_4/nrow(NBA))
SSE_4
RMSE_4
```

```{r}
# Read in test set
NBA_test = read.csv("NBA_test.csv")

# Make predictions on test set
PointsPredictions = predict(PointsReg4, newdata=NBA_test)

# Compute out-of-sample R^2
SSE = sum((PointsPredictions - NBA_test$PTS)^2)
SST = sum((mean(NBA$PTS) - NBA_test$PTS)^2)
R2 = 1 - SSE/SST
R2

# Compute the RMSE
RMSE = sqrt(SSE/nrow(NBA_test))
RMSE
```


